---
title: "Real Data Finally"
author: "Benyomin Hagalili"
date: "9/2/2016"
output: html_document
#output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Add another year of data to the previous graph  
Create a boxplot to illustrate the range of Incomes 
as it varies by city
[code gist](https://gist.github.com/paynito/14fb904b9f589e4c17aedde049fd16f1)
```{r create data}
# duplicate 1 # to make a mode
nyc_income1984 <- c(4,246,24,34,234,553,34)
dfw_income1984 <- c(344, 423,522, 234, 522, 255)
okc_income1984 <-c(234,234,236,99,184,301)
      # pad any short columns with NA
max.len = max(length(nyc_income1984),length(dfw_income1984),length(okc_income1984))
nyc1984 <-c(nyc_income1984, rep(NA, max.len-length(nyc_income1984)))
dfw1984 <-c(dfw_income1984, rep(NA, max.len-length(dfw_income1984)))
okc1984 <-c(okc_income1984,rep(NA,  max.len-length(okc_income1984)))
nyc1985 <- nyc1984*1.4
dfw1985 <- dfw1984+50
okc1985 <- okc1984*.7
#create data for a third year
nyc1986 <- nyc1984*.8
dfw1986 <- dfw1984*1.2
okc1986 <- okc1984*1.3
location <- c('NYC','DFW','OKC')
cityName <- tolower(location)
# year = c(rep(1984,3),rep(1985,3),rep(1986,3))
year<- c(1984,1984,1984,1985,1985,1985,1986,1986,1986)
# year<- c(1984,1985,1986)
library(reshape2)
                                        #combine the data
all.data = data.frame(rbind
                           (nyc1984,dfw1984,okc1984,
                             nyc1985,dfw1985,okc1985,
                             nyc1986,dfw1986,okc1986)
                      )
# this seems to end up with strings, not pointers to lists
#y1984 <-paste0(cityName, 1984)
#y1985 <- paste0(cityName,1985)
#y1986<-paste0(cityName,  1986)
#all.data = data.frame(rbind(y1984,y1985,y1986))
##
all.data$location <- location
all.data$year     <- year
stacked.data = melt(all.data,id =c('location','year'))
head(stacked.data)
incomeByCity <- stacked.data[,-3]
boxplots.triple = boxplot(value~location+year,
                          data = incomeByCity,
                          ylab = "$/month",
                          las = 2,
                        #  at = c(1,1.8,2.6,6,6.8,7.6),
                        #  names = location,
                        #  xaxt = 'n',
                          ylim=c(0,800),
                          col = c("sienna","palevioletred1","royalblue2")
                          )
# jpeg('groupedBoxPlots3.jpg')
     boxplots.triple
    # text(c(1,1.8,2.6,6,6.8,7.6),c(6,8,1,17.5,20,15.5),location)
    # axis(side=1, at=c(1.8,6.8,10.5), labels=c(1984,1985,1986))
     title('Individual Income in 3 cities, 1984-86')       
                                        # dev.off()

library(foreign)
library(Hmisc)
require(ggplot2)
require(magrittr)
# load rename function from plyr
require(plyr)
# load revised functions from dplyr
require(dplyr)
library(tidyr)
file456 <- spss.get("/Users/AbuDavid/school/census/f456/f456ind.por", use.value.labels=TRUE)
subSetInd2014<-dplyr::select(file456,YEARSUR,WPLDIST,USSRBORN,INCSAL)
#glimpse(subSetInd2014)
#utils::View(subSetInd2014)
#ggplot(subSetInd2014, aes(x=YEARSUR,y=INCSAL))+geom_boxplot()
##subset to only 3 cities, easier to copy DFW boxplot code
subCitiesInd2014 <- subSetInd2014 %>% dplyr::filter(WPLDIST=="Center"    |
                                                    WPLDIST=="Tel-Aviv"  |
                                                    WPLDIST=="Gush Dan")
##check the subset
utils::View(incomeByCity)#
#utils::View(subCitiesInd2014)
```

Text section: why is it getting evaluated now?
How does your NYC/DFW data compare to the Israel Data?

|   NYC     | Israel           | 
| ------------- |:-------------:| 
|   location    | WPLDIST     |   
| year          | YEARSUR    |    
| value         | INCSAL     |   


```{r plot4}
 # remove USSRBORN column
subCities2014 <- subCitiesInd2014 %>% select(-USSRBORN)
#utils::View(subCities2014)

# reorder by column name
# data <- data[c("A", "B", "C")]
subCitiesIND2014 <- subCitiesInd2014[c("WPLDIST","YEARSUR","INCSAL")]
# utils::View(subCitiesIND2014)
# rename to match NYC namespace
 Israel2014 <-subCitiesIND2014 %>%rename(c('WPLDIST' = 'location',
                                          'YEARSUR' = 'year',
                                           'INCSAL' = 'value'))
                                        #year = YEARSUR,value = INCSAL)
                                        ## comment this section until after
                                        ##reshape2::melt  - or - tidyr::gather
                       ##                  head(Israel2014)

                                       
                       ##                  # create fake data for 2 more years
                       ##                  Israel2015 <-Israel2014
                       ##                  Israel2016 <-Israel2014
                       ##                  # give everyone a raise in salary
                                     
                       ##                  Israel2015$value <- Israel2015$value * 1.2
                                       
                       ##                  Israel2015$year <- 2015
                       ##                  Israel2016$value <- Israel2015$value * 1.2
                       ##                  Israel2016$year <- 2016
                       ##                  # create stacked data frame
                       ## #         all.Israel= data.frame(rbind
                       ##     #     (Israel2014,Israel2015,Israel2016)) 
                       ##                  #        rbind(head(Israel2014),head(Israel2015),head(Israel2016)
                       ##                  all.Israel <-rbind(Israel2014,Israel2015,Israel2016)
                       ##                  head(all.Israel)
                       ##                  head(incomeByCity)
                       ##                  ## do the city's have to rotate through the list, or just have
                       ##                  ## the same sample size each??
                       ##                  ## get length
                                         
                                        length(Israel2014$location)
                                       
                                        count(Israel2014$location=="Center")
                                        count(Israel2014$location=="Gush Dan")
                                        count(Israel2014$location=="Tel-Aviv")
## previous code used
##   all.data = data.frame(rbind
##                            (nyc1984,dfw1984,okc1984,
##                              nyc1985,dfw1985,okc1985,
##                              nyc1986,dfw1986,okc1986))
## all.data$location <- location
## all.data$year     <- year
## stacked.data = melt(all.data,id =c('location','year'))

    ```
* Some words after final code block.
** The opposite of tidyr::gather is spread.
** The opposite of reshape2::melt is separate() or colsplit().
```{r plot5}
#tidyr::spread(Israel2014)
Israel2014 %>% filter(location=="Center")

```
